---
title: "coursera_pml"
author: "peterkratky"
date: "8 júla 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
```

## Coursera Project

This is a guide of how I predicted the excercise execusion classes based on sensor data.

### Preprocessing data

First, I loaded the training data and excluded rows with `new_window` variable equal to `yes`.
This is because the test dataset contains only rows with `new_window` equal to `no`.

```{r, include=TRUE}
data = read.csv("pml-training.csv")
data = subset(data, new_window == "no")
print(paste("rows: ", nrow(data), ", cols: ", ncol(data)))
```

Further, I removed all columns which contained at least one NA.

```{r, warning = FALSE}
cols = c()
for (i in 8:(ncol(data)-1)) {
  if (!is.na(mean(data[,i])))
    cols = c(cols, i)
}
data = data[, c(cols,ncol(data))]
print(paste("rows: ", nrow(data), ", cols: ", ncol(data)))
```

Finally, I removed highly correlated attributes.

```{r, warning = FALSE}
correlationMatrix <- cor(data[,8:(ncol(data)-1)], use = "everything")
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.5)
data = data[, -highlyCorrelated]
print(paste("rows: ", nrow(data), ", cols: ", ncol(data)))
```

```{r, warning = FALSE}
print(colnames(data))
```

### Splitting dataset

I splitted the training dataset to training (70%) and testing (30%) part.
Training is for building the model and tuning parameters.
Testing is to demonstrate the out of sample performance.

```{r, warning = FALSE}
train = createDataPartition(y=data$classe, p=0.7, list=FALSE)
training = data[train,]
testing = data[-train,]

```


### Training the model

I trained a clasifier and using 10-fold cross validation I checked its performance and tuned parameters.
Also I used Box-Cox transformation because some of the variables look rather skewed.

```{r, message = FALSE, warning=FALSE}
library(ggplot2)
library(reshape2)
dataForPlot <- melt(training[,5:8])
ggplot(dataForPlot, aes(value)) + 
  geom_bar() + 
  facet_wrap(~ variable, scales="free", ncol=4)
```

As the outcome variable is discrete, I used decision tree for classification. It had poor accuracy.
I decided to use Random Forrest classifier as it combines multiple decision trees

```{r, message = FALSE, warning=FALSE}
set.seed(123)
control <- trainControl(method="repeatedcv", number=10, repeats=1)
model <- train(classe~., data=training, method="rf", preProcess=c("BoxCox"), trControl=control)
print(model)
```


### Testing the model

I tested the model on the testing part of the dataset.
The accuracy is around 0.97 and I can conclude that this is out of sample acuracy.

```{r, message = FALSE, warning=FALSE}
pr = predict(model, testing)
m = confusionMatrix(pr, testing$classe)
print(m)
```

### Getting predictions for the quiz

I had to apply the preprocessing to testing dataset.

```{r, message = FALSE, warning=FALSE}
test = read.csv("pml-testing.csv")
test = test[, c(cols)]
test = test[, -highlyCorrelated]
prTest = predict(model, test)
```





